---
title: "What is civil war?"
menu:
  assignment:
    parent: Assignments
    weight: 2
---



```{r setup, include=FALSE}
# set knit options
knitr::opts_chunk$set(
  fig.width=9, 
  fig.height=5, 
  fig.retina=3,
  fig.align="center",
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  eval = FALSE,
  message = FALSE, 
  warning = FALSE
)

```

## Instructions



- For the data, right-click, "save/download file as..."
- Open the data
- Complete all of the tasks associated with that data and keep track of your answers / images
- **Once you are done**, go on Canvas and answer the questions related to the assignment
    - the Canvas assignment is **timed**; if you open it prior to completing the tasks you will run out of time





Skills used: 

- Sorting data
- Counting categories
- Cross-tabs
- Creating variables

[My awkward Google Sheets video tutorials are here](https://www.dropbox.com/sh/rsj9qm67se619pm/AAD1wPt84O8yaIfSPcwN9edua?dl=0)



::: {.callout-warning}
You must complete all of the tasks here BEFORE opening the Canvas quiz; otherwise, you will run out of time!
:::



## Data


You're going to work with the data from the Kalyvas and Balcells article we read. The dataset contains information on major civil wars around the world for a select time period. 


- The data: <i class="fas fa-download"></i> [Kalyvas and Balcells](https://www.dropbox.com/scl/fi/rpqb5sryxp5wqeivjqui2/balcells-kalyvas.csv?rlkey=jysyuzvj4xvhv6yzpn1jiepdq&dl=1)


```{r, warning=FALSE, echo=FALSE, message=FALSE, results='asis', eval=TRUE}
library(tidyverse)
# dictionary
tribble(~Variable, ~Description, 
        "countryname", "Country where civil war takes place",
        "year/month_start", "Year/month war began",
        "year/month_end", "Year/month war ended (if at all)",
        "conflict_name", "Name of conflict",
        "region", "World region",
        "gdp_capita", "GDP per capita (logged)",
        "rough_terrain", "Percent of country covered in rough terrain",
        "military_personnel_percapita", "Military personnel, in thousands",
        "marxist", "Is rebellion Marxist?",
        "technologyrebellion", "What kind of civil war is it?",
        "ethnicwar", "Is this an ethnic civil war?") %>% 
  pander::pandoc.table()

```



## Questions


Using the dataset, answer the following questions:


1. Choose a civil war of each **type** in the dataset and look it up on Wikipedia. Provide a non-technical, 2 sentence description of each war. (total = 3 wars, 2 sentences each). (2 points)



2. Sort the data. Which country had the largest army per capita involved in a civil war? (1 point)

```{r}
library(tidyverse)
df = read_csv("https://www.dropbox.com/scl/fi/rpqb5sryxp5wqeivjqui2/balcells-kalyvas.csv?rlkey=jysyuzvj4xvhv6yzpn1jiepdq&dl=1")
df |> 
  arrange(desc(military_personnel_percapita))
```



3. Count the number of civil wars by region. Which region has the highest number of recorded civil wars? (1 point)

```{r, eval = FALSE}
df$region |> table()
```


4. Cross-tabulate the number of civil wars across `warfare technology` and whether or not it's an `ethnic` conflict. How many civil wars are both "conventional" and ethnically motivated? (1 point)


```{r, eval = FALSE}
table(df$technologyrebellion, df$ethnicwar)
```




5. Create a variable that tells you how long each civil war has lasted, using only the years column (year ended - year started. Note: don't worry about civil wars that don't have a start or end date, you can ignore this issue). In which country did the *THIRD* longest lasting civil war take place? (1 point)


```{r}
df |> 
  mutate(duration = year_end - year_start) |> 
  arrange(desc(duration))
```


6. Filter this analysis down so you're only looking at Latin America. What is the SECOND longest conflict in this region? (1 point)

```{r}
df |> 
  filter(region == "Lamerica") |> 
  mutate(duration = year_end - year_start) |> 
  arrange(desc(duration))
```


7. We we will frequently talk about civil wars "starting" and "ending" in this class, but figuring out exactly *when* this happens is tricky. If you were to design a rule for deciding when a civil war has **begun**, what would that rule be? What would you have to observe, exactly, to say that a war has broken out? (3 points)

    A) describe your rule, make it as specific as possible. For example: "*We can say that a civil war has begun when X, Y, and Z happens...*"
    
    B) Identify some "edge cases" that will cause trouble for that rule. 4-5 sentences. 