---
title: "Counterinsurgency"
menu:
  assignment:
    parent: Assignments
    weight: 4
---




```{r setup, include=FALSE}
# set knit options
knitr::opts_chunk$set(
  fig.width=9, 
  fig.height=5, 
  fig.retina=3,
  fig.align="center",
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  eval = FALSE,
  message = FALSE, 
  warning = FALSE
)

```

## Instructions



- For the data, right-click, "save/download file as..."
- Open the data
- Complete all of the tasks associated with that data and keep track of your answers / images
- **Once you are done**, go on Canvas and answer the questions related to the assignment
    - the Canvas assignment is **timed**; if you open it prior to completing the tasks you will run out of time





Skills used:

- Creating new variables
- Counting categories


[My awkward Google Sheets video tutorials are here](https://www.dropbox.com/sh/rsj9qm67se619pm/AAD1wPt84O8yaIfSPcwN9edua?dl=0)



::: {.callout-warning}
You must complete all of the tasks here BEFORE opening the Canvas quiz; otherwise, you will run out of time!
:::



## Task 1: Combat Obscura


1. Connect the film we watched in class, *Combat Obscura*, to two of the readings we've covered in class so far. What parallels do you see? What is striking or interesting to you about the film? ~250 words. (5 points).

::: {.callout-note}
If you weren't able to attend lecture, or would like to review the film, it is available on Kanopy through UC Davis.
:::



## Task 2: US casualties



Let's look at this data on American casualties in the Iraq and Afghanistan wars (variable dictionary below). Each row here is a soldier who died in one of these two wars.

- Download this: <i class="fas fa-download"></i> [iCasualties data](https://www.dropbox.com/scl/fi/n5ha5rnqzoksm8yviovy0/casualties.csv?rlkey=w0syedcrxd8zyr45czh9a6ep8&dl=1)


2. Pick a row at random — look up the soldier online and see what you can find. What happened to this person? Who were they? (2 point)

```{r}
library(tidyverse)
df = read_csv("https://www.dropbox.com/scl/fi/n5ha5rnqzoksm8yviovy0/casualties.csv?rlkey=w0syedcrxd8zyr45czh9a6ep8&dl=1")
```



3. About what percent of casualties are the result of non-hostile actions (things like accidents, friendly fire, etc.)? Look at the `source` variable. (1 point)


```{r}
df |> group_by(source) |>
  tally() |>
  mutate(percent = n / sum(n) * 100) |>
  filter(source == "non-hostile")
```



4. What was the most common `cause` of death for US soldiers? (1 point)


```{r}
df |> group_by(cause) |>
  tally() |>
  arrange(desc(n))
```



8. Figure out how to use the `MONTH` function in excel / sheets to create a new column that tells you in what month the person died, based on the date of death variable. In what month of the year were the most American casualties registered?


```{r}
df |>
  mutate(month = month(date, label = TRUE)) |>
  group_by(month) |>
  tally() |>
  arrange(desc(n)) |>
  filter(n == max(n))
```


```{r, echo=FALSE,message=FALSE,warning=FALSE,results="asis",eval=TRUE}
library(tidyverse)
library(pander)
tribble(~Variable, ~Description,
        "date", "Date of death",
        "name", "Name of soldier",
        "rank", "Rank of soldier",
        "nationality", "Nationality of soldier",
        "branch", "Branch of military",
        "age", "Age at death",
        "country", "Country (Iraq or Afghanistan)",
        "province", "Province where soldier died",
        "where", "City/locale where soldier died",
        "source", "Source of death (hostile or non-hostile)",
        "cause", "Cause of death",
        "state", "Home state of soldier",
        "city", "Home city/town of soldier",
        "state_pop_2010", "Population in state (2010)") %>%
  pandoc.table()
```