---
title: "The war of the flea"
menu:
  assignment:
    parent: Assignments
    weight: 5
---




```{r setup, include=FALSE}
# set knit options
knitr::opts_chunk$set(
  fig.width=9, 
  fig.height=5, 
  fig.retina=3,
  fig.align="center",
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  eval = FALSE,
  message = FALSE, 
  warning = FALSE
)

```

## Instructions



- For the data, right-click, "save/download file as..."
- Open the data
- Complete all of the tasks associated with that data and keep track of your answers / images
- **Once you are done**, go on Canvas and answer the questions related to the assignment
    - the Canvas assignment is **timed**; if you open it prior to completing the tasks you will run out of time





Skills used: 

- grouped summary counts
- grouped summary statistics


[My awkward Google Sheets video tutorials are here](https://www.dropbox.com/sh/rsj9qm67se619pm/AAD1wPt84O8yaIfSPcwN9edua?dl=0)



::: {.callout-warning}
You must complete all of the tasks here BEFORE opening the Canvas quiz; otherwise, you will run out of time!
:::



## The Vietnam War


Here, you'll be working with hamlet level data on territorial control, selective violence, and indiscriminate violence during the Vietnam War. Each row is a hamlet in a given month in the year 1969. Using the data (variable dictionary below), answer the following questions. 


- Download this: <i class="fas fa-download"></i> [Vietnam HES data](https://www.dropbox.com/scl/fi/iq9jmh9gycjkal58wyr2c/vietnam.csv?rlkey=imdm954j7n9c6j01rvqlxv89n&dl=1)



1. What was the Strategic Hamlet Program during the Vietnam War? Look it up, describe in your own words, four to five sentences. Tell us what the program *did* and what its intended *objective* was. (4 points)


```{r}
library(tidyverse)
df = read_csv("https://www.dropbox.com/scl/fi/iq9jmh9gycjkal58wyr2c/vietnam.csv?rlkey=imdm954j7n9c6j01rvqlxv89n&dl=1")
```



2. HOW MANY bomb sorties were launched, in total, against the most attacked hamlet in Vietnam? (1 point)


```{r}
df |> 
  group_by(hamlet_id) |> 
  summarise(bombed = sum(bombed)) |> 
  filter(bombed == max(bombed, na.rm = TRUE))
```



3. Calculate the PERCENT of observations in each control category. Overall, who had more control, the rebels, or the government? (1 point)

```{r}
df |> group_by(control) |> 
  tally() |> 
  mutate(percent = n / sum(n) * 100)
# the state
```


4. About HOW MANY MORE bomb sorties, in total, were aimed at places that had full rebel control than places where the government had full control? (1 point)


```{r}
df |> 
  group_by(control) |> 
  summarise(bombed = sum(bombed, na.rm = TRUE))
9426 - 529
```



5. How common was the use of selective violence during the war? Take the average of selective violence across areas of control (note: remember what the average of a 0/1 variable tells you!). Where was it most or least common? How common would you say it was overall? Reference specific quantities in your answer. (2 points)

```{r}
df |> 
  group_by(control) |> 
  summarise(selective_terror = mean(selective_terror, na.rm = TRUE)) |> 
  mutate(selective_percent = selective_terror * 100)
# overall, infrequent: in the highest violece category, only 7.8% of observations experienced selective terror. Least common was in places of full state control (about .1%), most common was in places of majority (but not full) rebel control (7.8%). 
```



6. Do anything else you want with the data. Describe what you did and what you found. (2 points)







```{r, echo=FALSE,message=FALSE,warning=FALSE,results="asis",eval=TRUE}
library(tidyverse)
library(pander)
tribble(~Variable, ~Description, 
        "hamlet_id", "9-digit HES identifier",
        "month", "Month",
        "selective_terror", "Did Vietcong use selective violence? (1 = yes, 0 = no)",
        "control", "Who controlled hamlet?\n(1 = full state, 2 = partial state, 3 = contested, 4 = partial rebel, 5 = full rebel)",
        "bombed", "How many bombing sorties in hamlet?",
        "urban", "Is hamlet mostly urban?\n(1 = yes, 0 = no)",
        "pop", "Hamlet population",
        "border", "Distance to closest international border",
        "dev_index", "Development index (note: this is an index so the exact numbers are not interpretable. Lower numbers = less developed, larger numbers = more developed.)",
        "buddhist", "Predominantly Buddhist hamlet?\n(1 = yes, 0 = no)",
        "vietnamese", "Predominantly Vietnamese hamlet?\n(1 = yes, 0 = no)") %>% 
  pandoc.table()
```




