---
title: "Displacement"
menu:
  assignment:
    parent: Assignments
    weight: 6
---




```{r setup, include=FALSE}
# set knit options
knitr::opts_chunk$set(
  fig.width=9, 
  fig.height=5, 
  fig.retina=3,
  fig.align="center",
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  eval = FALSE,
  message = FALSE, 
  warning = FALSE
)

```

## Instructions



- For the data, right-click, "save/download file as..."
- Open the data
- Complete all of the tasks associated with that data and keep track of your answers / images
- **Once you are done**, go on Canvas and answer the questions related to the assignment
    - the Canvas assignment is **timed**; if you open it prior to completing the tasks you will run out of time





Skills used: 

- Summary statistics
- Plotting
- Creating variables


[My awkward Google Sheets video tutorials are here](https://www.dropbox.com/sh/rsj9qm67se619pm/AAD1wPt84O8yaIfSPcwN9edua?dl=0)



::: {.callout-warning}
You must complete all of the tasks here BEFORE opening the Canvas quiz; otherwise, you will run out of time!
:::


## Displacement around the world


You are going to work with data on the number of refugees around the world from the United Nations High Commissioner for Refugees. We are also going to look at data on deportations from the European Union (source: Eurostat).

- <i class="fas fa-download"></i> [UNHR refugee numbers (UNHR OVERALL)](https://www.dropbox.com/scl/fi/r3470e5fci18zj9rt8kwj/unhr-overall.csv?rlkey=4cju5geci2noaae2bi8j1dici&dl=1)

- <i class="fas fa-download"></i> [Refugee country of origin (ORIGIN)](https://www.dropbox.com/scl/fi/gcrb3i1rgx7372ohn3axc/origin.csv?rlkey=loc6kl6074py8hpx97jl1zhpv&dl=1)

- <i class="fas fa-download"></i> [Refugee country of destination (HOST)](https://www.dropbox.com/scl/fi/b2ydio9d1zs7qtavh1aut/host.csv?rlkey=wnndmfnothxwaddw50fizubha&dl=1)

- <i class="fas fa-download"></i> [European Union deportation figures (EU DEPORT)](https://www.dropbox.com/scl/fi/il416tqnnf2j5xu66gciq/eu_deport.csv?rlkey=gluojb8tua0g0ma9g2qauf85a&dl=1)


```{r, eval = TRUE}
# clean up names
labs = tibble::tibble(data = c("refugees_under_unhc_rs_mandate", 
                       "asylum_seekers", 
                       "id_ps_of_concern_to_unhcr",
                       "venezuelans_displaced_abroad", 
                       "palestine"), 
              meaning = c("Refugees under UNHCR mandate", 
                      "Asylum-seekers", 
                      "Internally displaced people of concern to UNHCR", 
                      "Venezuelans displaced abroad", 
                      "Palestinian refugees"))

kableExtra::kbl(labs, caption = "Description of columns, types of refugees")
```


```{r, eval = FALSE}
library(tidyverse)


# read palestinians
palestine = read_csv("/Users/juan/Dropbox/papers/wb-displace-review/data/palestine.csv") %>% 
  janitor::clean_names() %>% 
  select(year, palestine = total)

# read overall unhr data
overall = read_csv("/Users/juan/Dropbox/papers/wb-displace-review/data/unhr-overall.csv") %>% 
  janitor::clean_names() %>% 
  select(year, refugees_under_unhc_rs_mandate:venezuelans_displaced_abroad) %>% 
  # replace venezulan NAs with 0
  mutate(venezuelans_displaced_abroad = 
           replace_na(venezuelans_displaced_abroad, 0)) %>% 
  left_join(palestine, by = "year")





# spit out csv
write_csv(overall, file = "../files/datasets/unhr-overall.csv")


country = read_csv("/Users/juan/Dropbox/papers/wb-displace-review/data/unhr-country.csv") %>% 
  janitor::clean_names() %>% 
  select(year, country_of_origin, 
         refugees_under_unhc_rs_mandate, asylum_seekers, 
         id_ps_of_concern_to_unhcr)

write_csv(country, file = "../files/datasets/origin.csv")

host = read_csv("/Users/juan/Dropbox/papers/wb-displace-review/data/unhr-host.csv") %>% 
  janitor::clean_names() %>% 
  select(year, country_of_asylum, 
         refugees_under_unhc_rs_mandate, asylum_seekers, 
         id_ps_of_concern_to_unhcr)

write_csv(host, file = "../files/datasets/host.csv")

# EU deportation data
eu = readxl::read_excel("/Users/juan/Dropbox/papers/wb-displace-review/data/EU_deportations.xlsx", sheet = "Sheet 1", skip = 10) %>% 
  # clean up the data
  select(-contains("...")) %>% 
  rename(country = TIME) %>% 
  slice(-c(1:3)) |> 
  mutate(across(-country, as.numeric)) |> 
  mutate(country = ifelse(str_detect(country, "Germany"), "Germany", country))

write_csv(eu, file = "../files/datasets/eu_deport.csv")

# pulled this table from the DHS yearbook statistics: 
# https://www.dhs.gov/immigration-statistics/yearbook/2019/table39
us = tibble::tribble(
   ~year, ~removals, ~returns,
   2019L,     359885,    171445,
   2018L,     328716,    159940,
   2017L,     287093,    100695,
  2016L,     331717,    106458,
   2015L,     325328,    129661,
   2014L,     405090,    163843,
   2013L,     432228,    178976,
   2012L,     415607,    231107,
   2011L,     390423,    322047,
   2010L,     382449,    471798,
   2009L,     379739,    582567,
   2008L,     359795,    811263,
   2007L,     319382,    891390,
   2006L,     280974,   1043381,
   2005L,     246431,   1096920,
   2004L,     240665,   1166576,
   2003L,     211098,    945294,
   2002L,     165168,   1012116,
   2001L,     189026,   1349371,
   2000L,     188467,   1675876
  ) %>% 
  mutate(country = "United States")


write_csv(us, file = "../files/datasets/us_deport.csv")

```




1. Look at `UNHCR OVERALL`. How many TOTAL refugees exist in the world in the year 2020? (1 point)


```{r}
library(tidyverse)
df = read_csv("https://www.dropbox.com/scl/fi/r3470e5fci18zj9rt8kwj/unhr-overall.csv?rlkey=4cju5geci2noaae2bi8j1dici&dl=1")
df |> 
  rowwise() %>% 
  mutate(sum = sum(c_across(-year))) |> 
  ungroup() |> 
  filter(year == max(year)) |> 
  pull(sum)
```



2a. Look at `UNHCR OVERALL`. Make a time series plot with year on the x-axis and the trends for each type of refugee category on the y-axis. Each "type" of refugee in the data should be its own line. Save the image to upload. (1 point)


```{r}
df |> 
  pivot_longer(-year) |> 
  ggplot(aes(x = year, y = value, color = name)) + 
  geom_line()
```



2b. In a few sentences, describe the trends you observe in the image. (1 point)




3. Create a new variable that tells you what percent of all refugees are IDPs. What is the first year in which IDPs constitute more than 50% of all refugees? (1 point)


```{r}
df |> 
  mutate(total = refugees_under_unhc_rs_mandate + asylum_seekers + id_ps_of_concern_to_unhcr + venezuelans_displaced_abroad + palestine,
         idps_pct = id_ps_of_concern_to_unhcr / total * 100) |> 
  filter(idps_pct > 50) |> 
  filter(year == min(year))
```


4. Look at `HOST`. For the latest year for which there is data, what country is the second largest recipient of asylum seekers? (1 point)

```{r}
df = read_csv("https://www.dropbox.com/scl/fi/b2ydio9d1zs7qtavh1aut/host.csv?rlkey=wnndmfnothxwaddw50fizubha&dl=1")
df |> 
  filter(year == max(year)) |> 
  slice_max(order_by = asylum_seekers, n = 5) |> 
  slice(2)
```



5. Do some googling. *Why* is this country receiving so many asylum applicants? (1 point)

```{r}
# huge outflow of Venezuelans seeking asylum in Peru in 2020
```



6. Look at the data on deportations from the European Union. Create a variable that tells you how much deportations grew or shrunk from 2011 to 2020 for each country. What is the AVERAGE change in deportation from 2011 to 2020 across the European Union? (1 point)

```{r}
df = read_csv("https://www.dropbox.com/scl/fi/il416tqnnf2j5xu66gciq/eu_deport.csv?rlkey=gluojb8tua0g0ma9g2qauf85a&dl=1")
df |> 
  mutate(diff = `2020` - `2011`) |> 
  summarise(mean(diff, na.rm = TRUE))
```







## Stories of displacement

7. You can find many stories on the internet of people who have been displaced or are living as refugees in different parts of the world. Find one of these stories and, in 5-6 sentences, tell us about the person, what happened to them, and where they are now. Include the url to the story in your entry. (4 points)